pema application

### 2022-12-01 paper/manuscript.Rmd
final corrections

---
title             : "Select relevant moderators using Bayesian regularized meta-regression"
shorttitle        : "BAYESIAN REGULARIZED META-REGRESSION"

@@

One approach to is to select variables on theoretical grounds.
An important caveat is that theories that describe phenomena at the level of individual units of analysis do not necessarily generalize to the study level.
Using such theories for variable selection amounts to committing the ecological fallacy:
generalizing inferences across levels of analysis [@jargowskyEcologicalFallacy2004].
The implications of the ecological fallacy for interpreting the *results* of meta-regression are well-known [@bakerUnderstandingHeterogeneityMetaanalysis2009; @thompsonHowShouldMetaregression2002]:
For example, meta-regression may find a significant positive effect of average sample age on the effect size of a randomized controlled trial,
even if age is uncorrelated with treatment efficacy within each study.
Less well-known is that the same problem applies when using individual level theory to select study level moderators:
If theory states that an individual's age influences their susceptibility to treatment,
that does not imply that average sample age will be a relevant moderator of study-level treatment effect in meta-regression.
One rare example of study level theory is the *decline effect*:
effect sizes in any given tranche of the literature tend to diminish over time [@schoolerUnpublishedResultsHide2011].
When, by coincidence, a large effect is found, it initially draws attention from the research community.
Subsequent replications then find smaller effects due to regression to the mean.
Based on the decline effect, we might hypothesize "year of publication" to be a relevant moderator of study effect sizes.
At present, few such study level theories about the drivers of heterogeneity exist, and until they are developed,
theory has limited utility for variable selection.</span>
A further complication is that theoretically relevant variables might not be reported in many published papers,
which may be one reason why moderator analyses are rarely executed as planned [@thompsonHowShouldMetaregression2002]

@@

The *random effects* model, by contrast, assumes that,
in addition to sampling error, true effects vary for random reasons,
and thus follow a normal distribution with mean $\beta_0$ and variance $\tau^{2}$ [@hedgesFixedRandomeffectsModels1998].
This assumption is appropriate when studies are conceptually similar but differ in small random ways [@higginsReevaluationRandomeffectsMetaanalysis2009a]. 
The observed effect sizes are thus given by:
\begin{align}
T_i &= \beta_0 + \zeta_{i} + \epsilon_i\\
\text{where } \zeta_i &\sim N(0, \tau^2)\\
\text{and } \epsilon_i &\sim N(0, v_i)
\end{align}
In this model, the error term $\zeta_{i}$ represents between-studies heterogeneity, with variance $\tau^2$.
As in the fixed effect model,
studies with smaller sampling errors are assigned more weight.
<span id = "informationdistribution">In contrast to the fixed effect model, the random effects model assumes that all studies provide some information about the underlying distribution of true effect sizes.
Fixed effect weights would discount the information smaller studies provide about the scale of this distribution, which is represented by it's variance $\tau^2$.
To overcome this limitation, the weights are attenuated in proportion to the variance.
The random effects weights are thus given by $w_{i} =  \frac{1}{v_i + \hat{\tau}^2}$.</span>
Whereas sampling error is still treated as known,
the between-study heterogeneity $\tau^{2}$ must be estimated.
This estimate is represented by $\hat{\tau}^{2}$.

@@

As explained in Formula \@ref(eq:),
regularization penalizes all coefficients equally,
without regard for their scale.
If variables are on different scales,
this can lead to uneven penalization of coefficients in which
variables with smaller standard deviations are biased more strongly towards zero [@lee2015note].
If the scale of predictor $x$ is increased by a factor 10,
its regression coefficient is reduced by a factor 10,
bringing it closer to zero where it will be more affected by penalization.
Standardization is a widely used method for equalizing predictor scales [@gelmanScalingRegressionInputs2008a].
Standardization is a linear transformation that sets the mean of all predictors to 0 and their standard deviation to 1.
Like most other regularizing methods,
BRMA performs standardization by default [@tibshirani1996].
After parameters are estimated using standardized variables,
they can be restored to their original scales.
For the intercept, the transformation is:
$$
b_0 = b_{0Z} - \mathbf{b}_Z\frac{\bar{\mathbf{x}}}{\mathbf{s}_X}
$$
where $b_0$ is the intercept, $b_{0Z}$ is the intercept for the standardized predictors,
$\mathbf{\bar{x}}$ and $\mathbf{s}_{x}$ are the vectors of predictor means and variances,
and $\mathbf{b_Z}$ is the vector of regression coefficients for the standardized predictors.
The regression coefficients are returned to their original scale by applying:
$$
\mathbf{b}_x = \frac{\mathbf{b}_z}{\mathbf{s}_x}
$$

@@

Any iterative algorithm is susceptible to convergence problems.
In such cases, the BRMA algorithms provide warning messages,
but still return samples from the posterior.
We were thus able to use all iterations of the BRMA algorithms,
although some may have failed to converge, which would negatively impact BRMA's performance.
When the RMA algorithm fails to converge, however,
terminates with an error.
<!-- To handle this contingency, we automated some of the steps recommended [on the `metafor` website](https://www.metafor-project.org/doku.php/tips:convergence_problems_rma). -->
<!-- This means that, in case of convergence  -->
The RMA algorithm failed to converge in `r out$missing` replications,
all characterized by low number of cases ($k \leq 40$) and high effect sizes $\beta \geq .5$.
They were omitted from further analysis.

@@

Before interpreting the results, however,
it is important to assess model convergence.
If any indication of non-convergence is detected during estimation,
a warning will be printed.
The example returns the warning that there were 331 divergent transitions,
and suggests increasing the number of iteractions (increasing the argument `iter` beyond its default value of 2000).
Divergent transitions can result in biased estimates.
If the number of divergences is small and there are no further indications of non-convergence, however,
the posterior distribution is often good enough to safely interpret the results.
We can examine two parameter-specific indicators of convergence
by ascertaining that the number of "unique" samples from the posterior `n_eff` for each parameter is sufficiently high,
and that the different chains of the estimator have mixed properly, as indicated by `Rhat` close to 1.
The number of effective (independent) MCMC samples should be high relative to the total number of samples (in this case, 4000, as we used 2000 iterations on a dual-core processor).
If the effective sample size is less than 10% of the total,
there may be a problem - which is not the case here.
The `Rhat` is a version of the potential scale reduction factor,
which represents the ratio of between- and within-chain variance [@GelmanRubin1992].
If the chains mixed well, the `Rhat` should be close to 1.
Both `n_eff` and `Rhat` indicate convergence in this example.
Additional convergence diagnostics are obtained by running `check_hmc_diagnostics(as.stan(fit))`.
Convergence can also be assessed visually using the function `traceplot(as.stan(fit), pars = c("Intercept", "year"))`,
which provides trace plots for the MCMC draws.
If the model converged, the traces of the different chains should mix well (i.e., overlap) and look like "fat caterpillars".

@@

In the Introduction, researchers should substantiate the decision to explore heterogeneity
One valid reason is prima facie heterogeneity of the body of research [@higginsReevaluationRandomeffectsMetaanalysis2009a].
Another reason is the presence of theoretically relevant moderators [@thompsonHowShouldMetaregression2002].
Less convincing is the practice of exploring heterogeneity only when $\tau^2$ is significant,
for two reasons:
Firstly, because data-driven analysis decisions increase the risk of spurious findings [@zhangInferenceVariableSelection1992].
Secondly, because tests for heterogeneity are often underpowered when the number of studies is low, and overpowered when it is high, thus limiting their usefulness [@higginsQuantifyingHeterogeneityMetaanalysis2002].

@@

BRMA may not be the best solution for every situation.
Several trade-offs must be considered to decide what method is most appropriate.
Firstly, BRMA has higher predictive performance than RMA,
which implies that it is more suitable when a researcher intends to generalize beyond the sample at hand.
Conversely, RMA is be more suitable when the goal is to describe the sample at hand in an unbiased manner,
with less concern for generalizability to future studies.
Secondly, BRMA trades off higher specificity for lower sensitivity compared to RMA,
which suggests that it is more suitable when a researcher seeks to eliminate irrelevant moderators, at the cost of an increased Type II error rate.
RMA might be more suitable when the researcher seeks to identify relevant moderators, at the cost of a greater Type I error rate.
If many moderators are expected to be irrelevant,
then BRMA may thus be preferable.
Thirdly, there may be pragmatic reasons for preferring BRMA over RMA.
For example, if a data set is small, or the number of moderators is high relative to the number of cases,
RMA models may be empirically under-identified.
This can result in convergence problems.
In such cases, Bayesian estimation may converge on a solution where frequentist estimation does not [@kohliFittingLinearLinearPiecewise2015].
Similarly, BRMA may perform better in the presence of multicollinearity among predictors,
which can be examined using the function `vif()` in the R-package `metafor`.
Values exceeding 5 are cause for concern.
Multicollinearity increases the variance of regression coefficients.
BRMA may have an advantage here, because the regularizing priors restrict variance.
If multicollinearity is observed or suspected, BRMA might be preferred.

>>>>

---
title             : "Selecting relevant moderators with Bayesian regularized meta-regression"
shorttitle        : "BAYESIAN REGULARIZED META-REGRESSION"

@@

One approach is to select variables on theoretical grounds.
An important caveat is that theories that describe phenomena at the level of individual units of analysis do not necessarily generalize to the study level.
Using such theories for variable selection amounts to committing the ecological fallacy:
generalizing inferences across levels of analysis [@jargowskyEcologicalFallacy2004].
The implications of the ecological fallacy for interpreting the *results* of meta-regression are well-known [@bakerUnderstandingHeterogeneityMetaanalysis2009; @thompsonHowShouldMetaregression2002]:
For example, meta-regression may find a significant positive effect of average sample age on the effect size of a randomized controlled trial,
even if age is uncorrelated with treatment efficacy within each study.
Less well-known is that the same problem applies when using individual level theory to select study level moderators:
If theory states that an individual's age influences their susceptibility to treatment,
that does not imply that average sample age will be a relevant moderator of study-level treatment effect in meta-regression.
One rare example of study level theory is the *decline effect*:
effect sizes in any given tranche of the literature tend to diminish over time [@schoolerUnpublishedResultsHide2011].
When, by coincidence, a large effect is found, it initially draws attention from the research community.
Subsequent replications then find smaller effects due to regression to the mean.
Based on the decline effect, we might hypothesize "year of publication" to be a relevant moderator of study effect sizes.
At present, few such study level theories about the drivers of heterogeneity exist, and until they are developed,
theory has limited utility for variable selection.</span>
A further complication is that theoretically relevant variables might not be reported in many published papers,
which may be one reason why moderator analyses are rarely executed as planned [@thompsonHowShouldMetaregression2002]

@@

The *random effects* model, by contrast, assumes that,
in addition to sampling error, true effects vary for random reasons,
and thus follow a normal distribution with mean $\beta_0$ and variance $\tau^{2}$ [@hedgesFixedRandomeffectsModels1998].
This assumption is appropriate when studies are conceptually similar but differ in small random ways [@higginsReevaluationRandomeffectsMetaanalysis2009a]. 
The observed effect sizes are thus given by:
\begin{align}
T_i &= \beta_0 + \zeta_{i} + \epsilon_i\\
\text{where } \zeta_i &\sim N(0, \tau^2)\\
\text{and } \epsilon_i &\sim N(0, v_i)
\end{align}
In this model, the error term $\zeta_{i}$ represents between-studies heterogeneity, with variance $\tau^2$.
As in the fixed effect model,
studies with smaller sampling errors are assigned more weight.
<span id = "informationdistribution">In contrast to the fixed effect model, the random effects model assumes that all studies provide some information about the underlying distribution of true effect sizes.
Fixed effect weights would discount the information smaller studies provide about the scale of this distribution, which is represented by its variance $\tau^2$.
To overcome this limitation, the weights are attenuated in proportion to the variance.
The random effects weights are thus given by $w_{i} =  \frac{1}{v_i + \hat{\tau}^2}$.</span>
Whereas sampling error is still treated as known,
the between-study heterogeneity $\tau^{2}$ must be estimated.
This estimate is represented by $\hat{\tau}^{2}$.

@@

As explained earlier,
regularization penalizes all coefficients equally,
without regard for their scale.
If variables are on different scales,
this can lead to uneven penalization of coefficients in which
variables with smaller standard deviations are biased more strongly towards zero [@lee2015note].
If the scale of predictor $x$ is increased by a factor 10,
its regression coefficient is reduced by a factor 10,
bringing it closer to zero where it will be more affected by penalization.
Standardization is a widely used method for equalizing predictor scales [@gelmanScalingRegressionInputs2008a].
Standardization is a linear transformation that sets the mean of all predictors to 0 and their standard deviation to 1.
Like most other regularizing methods,
BRMA performs standardization by default [@tibshirani1996].
After parameters are estimated using standardized variables,
they can be restored to their original scales.
For the intercept, the transformation is:
$$
b_0 = b_{0Z} - \mathbf{b}_Z\frac{\bar{\mathbf{x}}}{\mathbf{s}_X}
$$
where $b_0$ is the intercept, $b_{0Z}$ is the intercept for the standardized predictors,
$\mathbf{\bar{x}}$ and $\mathbf{s}_{x}$ are the vectors of predictor means and variances,
and $\mathbf{b_Z}$ is the vector of regression coefficients for the standardized predictors.
The regression coefficients are returned to their original scale by applying:
$$
\mathbf{b}_x = \frac{\mathbf{b}_z}{\mathbf{s}_x}
$$

@@

Any iterative algorithm is susceptible to convergence problems.
In such cases, the BRMA algorithms provide warning messages,
but still return samples from the posterior.
We were thus able to use all iterations of the BRMA algorithms,
although some may have failed to converge, which would negatively impact BRMA's performance.
When the RMA algorithm fails to converge, however,
it terminates with an error.
<!-- To handle this contingency, we automated some of the steps recommended [on the `metafor` website](https://www.metafor-project.org/doku.php/tips:convergence_problems_rma). -->
<!-- This means that, in case of convergence  -->
The RMA algorithm failed to converge in `r out$missing` replications,
all characterized by low number of cases ($k \leq 40$) and high effect sizes $\beta \geq .5$.
They were omitted from further analysis.

@@

Before interpreting the results, however,
it is important to assess model convergence.
If any indication of non-convergence is detected during estimation,
a warning will be printed.
The example returns the warning that there were 331 divergent transitions,
and suggests increasing the number of iterctions (increasing the argument `iter` beyond its default value of 2000).
Divergent transitions can result in biased estimates.
If the number of divergences is small and there are no further indications of non-convergence, however,
the posterior distribution is often good enough to safely interpret the results.
We can examine two parameter-specific indicators of convergence
by ascertaining that the number of "unique" samples from the posterior `n_eff` for each parameter is sufficiently high,
and that the different chains of the estimator have mixed properly, as indicated by `Rhat` close to 1.
The number of effective (independent) MCMC samples should be high relative to the total number of samples (in this case, 4000, as we used 2000 iterations on a dual-core processor).
If the effective sample size is less than 10% of the total,
there may be a problem - which is not the case here.
The `Rhat` is a version of the potential scale reduction factor,
which represents the ratio of between- and within-chain variance [@GelmanRubin1992].
If the chains mixed well, the `Rhat` should be close to 1.
Both `n_eff` and `Rhat` indicate convergence in this example.
Additional convergence diagnostics are obtained by running `check_hmc_diagnostics(as.stan(fit))`.
Convergence can also be assessed visually using the function `traceplot(as.stan(fit), pars = c("Intercept", "year"))`,
which provides trace plots for the MCMC draws.
If the model converged, the traces of the different chains should mix well (i.e., overlap) and look like "fat caterpillars".

@@

In the Introduction, researchers should substantiate the decision to explore heterogeneity.
One valid reason is prima facie heterogeneity of the body of research [@higginsReevaluationRandomeffectsMetaanalysis2009a].
Another reason is the presence of theoretically relevant moderators [@thompsonHowShouldMetaregression2002].
Less convincing is the practice of exploring heterogeneity only when $\tau^2$ is significant,
for two reasons:
Firstly, because data-driven analysis decisions increase the risk of spurious findings [@zhangInferenceVariableSelection1992].
Secondly, because tests for heterogeneity are often underpowered when the number of studies is low, and overpowered when it is high, thus limiting their usefulness [@higginsQuantifyingHeterogeneityMetaanalysis2002].

@@

BRMA may not be the best solution for every situation.
Several trade-offs must be considered to decide what method is most appropriate.
Firstly, BRMA has higher predictive performance than RMA,
which implies that it is more suitable when a researcher intends to generalize beyond the sample at hand.
Conversely, RMA is more suitable when the goal is to describe the sample at hand in an unbiased manner,
with less concern for generalizability to future studies.
Secondly, BRMA trades off higher specificity for lower sensitivity compared to RMA,
which suggests that it is more suitable when a researcher seeks to eliminate irrelevant moderators, at the cost of an increased Type II error rate.
RMA might be more suitable when the researcher seeks to identify relevant moderators, at the cost of a greater Type I error rate.
If many moderators are expected to be irrelevant,
then BRMA may thus be preferable.
Thirdly, there may be pragmatic reasons for preferring BRMA over RMA.
For example, if a data set is small, or the number of moderators is high relative to the number of cases,
RMA models may be empirically under-identified.
This can result in convergence problems.
In such cases, Bayesian estimation may converge on a solution where frequentist estimation does not [@kohliFittingLinearLinearPiecewise2015].
Similarly, BRMA may perform better in the presence of multicollinearity among predictors,
which can be examined using the function `vif()` in the R-package `metafor`.
Values exceeding 5 are cause for concern.
Multicollinearity increases the variance of regression coefficients.
BRMA may have an advantage here, because the regularizing priors restrict variance.
If multicollinearity is observed or suspected, BRMA might be preferred.
